On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-01-22 MS.tex
Remove abbreviations: RGR, AGR, WD, Hmax

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across ontogenetic stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one ontogenetic stage category: seedlings and adults, respectively.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that a strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rate (RGR) among seedlings growing under favourable conditions. Seedling potential RGR has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into different components showed that the trait specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. From the 1990s a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents, latitudes, and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information. Using these studies, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

Recent theoretical work \citep{Falster:2011ii} allows us to formulate hypotheses about the expected relationship between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size. This model extends earlier efforts to model growth as a function of traits \citep{Lambers:1992bj,Cornelissen:1996hf,Wright:2000kw,Enquist:2007ek}. However, whereas earlier models focused solely on mass-based RGR and did not predict a change in effect of traits with size, the revised equations do allow for these effects, while also making predictions for both absolute and relative growth in any of mass (eq. \ref{eq:eq1}), height (eq. \ref{eq:eq2}), or diameter (eq. \ref{eq:eq4}).

@@

\subsection*{H2: The relationship between WD and growth rate is expected to be negative across all plant sizes}

Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between WD and growth may also strengthen in larger plants. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is non-significant \citep[see also][]{Enquist-1999}.

\subsection*{H3: The relationship between Hmax and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model eqs. (\ref{eq:eq2}-\ref{eq:eq4}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of  growth is used (Table \ref{tab:trade-offs}).

\subsection*{H4: The relationship between seed mass and RGR is negative in seedlings, but weakens with increasing size}

@@

A big challenge when searching literature databases is to identify suitable studies from the vast corpus available. Initially we were interested to know whether data existed for a variety of traits; however the number of potential search terms made systemic searching for all these terms in Google Scholar impractical. A preliminary investigation of search results for a variety of traits on the Web of Science platform suggested that searching only on SLA and WD -- two of the most widely measured traits -- would capture $\sim$80\% of search records that would be returned by searching for a wider range of traits (see Appendix \ref{app:literature-snapshot}). We therefore narrowed our searching to only cover these two main traits.

Both Web of Science and Google Scholar databases were searched for records linking SLA and WD to growth rate. On Web of Science, we searched using  search terms: (``trait'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), where ``trait'' was either (``leaf mass per area'' OR ``specific leaf area'' OR ``leaf construction cost'') or (``wood density'' OR``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'' OR ``SSD'' ). On Google scholar we searched using terms: ((``SLA'' OR ``LMA'' OR ``specific leaf area'' OR ``leaf mass area'' OR ``wood density'' OR ``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'') AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''). This keyword search in Google Scholar returned approximately 18900 papers. We examined all records returned from the Web of Science search (701 articles) and the first 750 items from Google Scholar when ranked by ``relevance''. All searches were conducted between March and December 2014.

In total we inspected the full text of 583 articles (see Appendix \ref{app:literature-list} for list). Individual studies had to report the following information to be included in our dataset: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size or stage of plants. Despite searching only on SLA and WD, we recorded coefficient correlations for a list of 19 traits in total: the five for which we had prior expectations, plus any of 14 others commonly used in the trait literature. Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or AGR measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). In some cases, we asked the authors for additional information about the specific correlation coefficient value when this was not already provided.

The 19 traits recorded (Fig. \ref{fig:fig1}) capture different components of plant ecological strategies. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource distribution to leaf \citep{Lambers:1992bj}.

@@

\subsubsection*{Measurements under 'ideal' conditions}\label{ideal-vs-entire-dataset}

The entire dataset included studies differing in the growth measurement used (RGR or AGR), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. However RGR and AGR are fundamentally different for a lot of ecologists. And the number of species used to perform the correlation, the experimental conditions, and the concordance between stages used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, across \underline{\textgreater} 10 species, under unstressed conditions, and where the source of the trait and growth is the same (expect for Hmax and seed mass, where a particular stage is by definition required to measure the trait). This dataset is called the "ideal" dataset. Then we compared results with those of a more inclusive dataset - the entire dataset- with both RGR and AGR, and all others criteria relaxed.

@@

\subsection*{Main effects under 'ideal' conditions}

Overall, results were consistent with predictions, although with limitations arising for some traits due to very unequal (and limited) numbers of observations between growth stages. We identified only 55 studies with 137 correlations reporting data measured under ``ideal'' conditions: unstressed growth, growth measured as RGR, traits and growth recorded in same stage, and correlation coefficient calculated across \underline{\textgreater} 10 species.

For SLA, the correlation between RGR and the trait changed significantly with plant size in the manner expected (Fig. \ref{fig:fig3}a). The correlation between RGR and SLA shifted from strongly positive at seedling stage ($\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero). The correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages. Model fit improved significantly when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3}a; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4}).

For WD, Hmax, and seed mass, correlations with RGR were also consistent with expectations but our ability to test all hypotheses was at times limited due to low sample size. Indeed, while meta-analysis are possible even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e CI often spanned zero). 
For WD, results were consistent with the prediction for negative correlations with RGR across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginaly improved when stage was incorporated (p.value = 0.049). Note that the CIs were not crossing zero for seedling and sapling stage despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistency in the coefficient reported for each stage across studies.

For Hmax, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Here, Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. 
While these results were consistent with expectation, we could not distinguish with another alternative: Hmax was correlated to growth for adult stage, but the correlation between Hmax and growth at sapling stage was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between Hmax and growth at small plant size.

As expected for seed mass, the correlation with RGR shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig \ref{fig:fig3}c, CIs overlapping zero). However, the correlation between seed mass and RGR at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). Again we could not distinguish between these alternatives. But unlike Hmax, the number of coefficient reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficient reported rather than to a much smaller numbers of replicates.

For Aarea, results were consistent with a positive correlation with RGR at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages; as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

We now report results from the entire dataset. Our aim is to determine if additional data alters or reinforce the above conclusions, by increasing replication. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different ontogenetic stage to growth data) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

Results from the entire dataset reinforced our conclusions for both SLA and WD, set apart alternative explanations for seed mass, and allowed us to compare correlations with growth along ontogenetic stages for Aarea. For SLA and WD, the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ''ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across ontogenetic stage. The results were consistent with the prediction for a positive correlation with RGR across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with RGR at seedling stage; as the correlation was still positive when 8 correlations measured under low light conditions were added. For Hmax, there were still too few replicates to conclude undoubtely on the correlation to be absent for smaller sizes.

We had hypotheses for seed mass and SLA, about a change in the correlation between trait and growth as AGR was used instead of RGR, or as mass was used instead of dimater and height. For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping). For SLA, the correlation was expected to shift from positive to non significant if AGRmass was used instead of RGRmass for seedling stage; here none of the data have been measured with AGRmass at seedling stage (Fig. \ref{fig:figA8}). 
% Why a change with WD when it is measured in mass? All WD seedling are measured in RGRmass, and there are some mass data for both other stages too.  
For other traits, effect size were generally consistent between RGR and AGR when both measurement were available. Some differences occurred but they were not easily attributed to any single cause. For instance for WD, the correlation at juvenile stage shifted from negative with RGR to non significant with AGR (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for AGR came from \citet{Augspurger:1984ct} and unlike others they used wood density values from adult plants together with growth from juveniles.

@@

In total, 103 articles published between 1983 and 2014, and containing 551 correlations linking growth rate to any of the 19 traits across a range of ontogenetic stages were identified. Data reported here were variable in multiple ways. Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. The number of species used for each correlation ranged from 2 to 300 species, with an average of 31. Given this variety, it was not surprising to find a wide range of $r$ reported for each trait. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}); suggesting confounding with these factors is not a likely explanation of the major results reported.

@@

For five prominent traits, the patterns observed were consistent with specific predictions from a mechanistic growth model. This suggests the mechanisms through which traits affect growth may be correctly understood in the model. For SLA the meta-analysis confirmed quite strongly the predicted influence of plant size on the trait-growth correlation. For WD, Aarea, seed mass and Hmax results were consistent with predictions, but coverage was limited and additional studies would be helpful for some trait-growth-stage relationships.

@@

For seedlings our study showed, as expected, a negative relationship between seed mass and RGR (in mass or height or volume) for small plants, that dissipated as plants grew larger. This negative relationship arises because seedlings from larger seeds begin life larger themselves, as has been extensively discussed in the literature \citep[reviewed by][]{Turnbull:2012ew}.

>>>>

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across ontogenetic stages. Specific leaf area was correlated with relative growth rate in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with relative growth rate only in one ontogenetic stage category: seedlings and adults, respectively.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that a strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rate  among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that the trait specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. From the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents, latitudes, and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood (or stem) tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum (or asymptotic height), and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information. Using these studies, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

Recent theoretical work \citep{Falster:2011ii} allows us to formulate hypotheses about the expected relationship between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size. This model extends earlier efforts to model growth as a function of traits \citep{Lambers:1992bj,Cornelissen:1996hf,Wright:2000kw,Enquist:2007ek}. However, whereas earlier models focused solely on mass-based relative growth rate and did not predict a change in effect of traits with size, the revised equations do allow for these effects, while also making predictions for both absolute and relative growth in any of mass (eq. \ref{eq:eq1}), height (eq. \ref{eq:eq2}), or diameter (eq. \ref{eq:eq4}).

@@

\subsection*{H2: The relationship between wood density and growth rate is expected to be negative across all plant sizes}

Cheaper stem construction (low wood density or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is non-significant \citep[see also][]{Enquist-1999}.

\subsection*{H3: The relationship between maximum height and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model eqs. (\ref{eq:eq2}-\ref{eq:eq4}). For seedlings, where all individuals are investing solely in growth, we expect no influence of maximum height on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their maximum height value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller maximum height are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of  growth is used (Table \ref{tab:trade-offs}).

\subsection*{H4: The relationship between seed mass and relative growth rate is negative in seedlings, but weakens with increasing size}

@@

A big challenge when searching literature databases is to identify suitable studies from the vast corpus available. Initially we were interested to know whether data existed for a variety of traits; however the number of potential search terms made systemic searching for all these terms in Google Scholar impractical. A preliminary investigation of search results for a variety of traits on the Web of Science platform suggested that searching only on SLA and wood density -- two of the most widely measured traits -- would capture $\sim$80\% of search records that would be returned by searching for a wider range of traits (see Appendix \ref{app:literature-snapshot}). We therefore narrowed our searching to only cover these two main traits.

Both Web of Science and Google Scholar databases were searched for records linking SLA and wood density to growth rate. On Web of Science, we searched using  search terms: (``trait'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), where ``trait'' was either (``leaf mass per area'' OR ``specific leaf area'' OR ``leaf construction cost'') or (``wood density'' OR``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'' OR ``SSD'' ). On Google scholar we searched using terms: ((``SLA'' OR ``LMA'' OR ``specific leaf area'' OR ``leaf mass area'' OR ``wood density'' OR ``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'') AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''). This keyword search in Google Scholar returned approximately 18900 papers. We examined all records returned from the Web of Science search (701 articles) and the first 750 items from Google Scholar when ranked by ``relevance''. All searches were conducted between March and December 2014.

In total we inspected the full text of 583 articles (see Appendix \ref{app:literature-list} for list). Individual studies had to report the following information to be included in our dataset: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size or stage of plants. Despite searching only on SLA and wood density, we recorded coefficient correlations for a list of 19 traits in total: the five for which we had prior expectations, plus any of 14 others commonly used in the trait literature. Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). In some cases, we asked the authors for additional information about the specific correlation coefficient value when this was not already provided.

The 19 traits recorded (Fig. \ref{fig:fig1}) capture different components of plant ecological strategies. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass and maximum height capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here wood density and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource distribution to leaf \citep{Lambers:1992bj}.

@@

\subsubsection*{Measurements under ``ideal'' conditions}\label{ideal-vs-entire-dataset}

The entire dataset included studies differing in the growth measurement used (relative of absolute), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. However relative growth rate and absolute growth rate are fundamentally different for a lot of ecologists. And the number of species used to perform the correlation, the experimental conditions, and the concordance between stages used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. So we adopted a two steps approach, first we were conservative by focusing exclusively on relative growth rate, across \underline{\textgreater} 10 species, under unstressed conditions, and where the source of the trait and growth is the same (expect for maximum height and seed mass, where a particular stage is by definition required to measure the trait). This dataset is called the ``ideal'' dataset. Then we compared results with those of a more inclusive dataset - the entire dataset- with both relative growth rate and absolute growth rate, and all others criteria relaxed.

@@

\subsection*{Main effects under ``ideal'' conditions}

Overall, results were consistent with predictions, although with limitations arising for some traits due to very unequal (and limited) numbers of observations between growth stages. We identified only 55 studies with 137 correlations reporting data measured under ``ideal'' conditions: unstressed growth, growth measured as relative growth rate, traits and growth recorded in same stage, and correlation coefficient calculated across \underline{\textgreater} 10 species.

For SLA, the correlation between relative growth rate and the trait changed significantly with plant size in the manner expected (Fig. \ref{fig:fig3}a). The correlation between relative growth rate and SLA shifted from strongly positive at seedling stage ($\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero). The correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages. Model fit improved significantly when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3}a; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4}).

For wood density, maximum height, and seed mass, correlations with relative growth rate were also consistent with expectations but our ability to test all hypotheses was at times limited due to low sample size. Indeed, while meta-analysis are possible even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e CI often spanned zero).
For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p.value = 0.049). Note that the CIs were not crossing zero for seedling and sapling stage despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistency in the coefficient reported for each stage across studies.

For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage.
While these results were consistent with expectation, we could not distinguish with another alternative: maximum height was correlated to growth for adult stage, but the correlation between maximum height and growth at sapling stage was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.

As expected for seed mass, the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig \ref{fig:fig3}c, CIs overlapping zero). However, the correlation between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). Again we could not distinguish between these alternatives. But unlike maximum height, the number of coefficient reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficient reported rather than to a much smaller numbers of replicates.

For Aarea, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages; as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

We now report results from the entire dataset. Our aim is to determine if additional data alters or reinforce the above conclusions, by increasing replication. This entire dataset included correlations measured with both absolute growth rate and relative growth rate (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different ontogenetic stage to growth data) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: Wood density-seedling, maximum height-sapling, maximum height-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

Results from the entire dataset reinforced our conclusions for both SLA and wood density, set apart alternative explanations for seed mass, and allowed us to compare correlations with growth along ontogenetic stages for Aarea. For SLA and wood density, the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ``ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across ontogenetic stage. The results were consistent with the prediction for a positive correlation with relative growth rate across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage; as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were still too few replicates to conclude undoubtedly on the correlation to be absent for smaller sizes.

We had hypotheses for seed mass and SLA, about a change in the correlation between trait and absolute growth rate was used instead of relative growth rate, or as mass was used instead of diameter and height. For seed mass, effects size differed significantly between relative growth rate and absolute growth rate at seedling stage; as expected the correlation with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI not overlapping). For SLA, the correlation was expected to shift from positive to non significant if growth in mass was measured on absolute instead of relative terms; however, here none of the data have been measured with absolute rates in the seedling stage (Fig. \ref{fig:figA8}).
For other traits, effect size were generally consistent between relative growth rate and absolute growth rate when both measurement were available. Some differences occurred but they were not easily attributed to any single cause. For instance for wood density, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct} and unlike others they used wood density values from adult plants together with growth from juveniles.

@@

In total, 103 articles published between 1983 and 2014, and containing 551 correlations linking growth rate to any of the 19 traits across a range of ontogenetic stages were identified. Data reported here were variable in multiple ways. Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. The number of species used for each correlation ranged from 2 to 300 species, with an average of 31. Given this variety, it was not surprising to find a wide range of $r$ reported for each trait. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for wood density (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}); suggesting confounding with these factors is not a likely explanation of the major results reported.

@@

For five prominent traits, the patterns observed were consistent with specific predictions from a mechanistic growth model. This suggests the mechanisms through which traits affect growth may be correctly understood in the model. For SLA the meta-analysis confirmed quite strongly the predicted influence of plant size on the trait-growth correlation. For wood density, Aarea, seed mass and maximum height results were consistent with predictions, but coverage was limited and additional studies would be helpful for some trait-growth-stage relationships.

@@

For seedlings our study showed, as expected, a negative relationship between seed mass and relative growth rate (in mass or height or volume) for small plants, that dissipated as plants grew larger. This negative relationship arises because seedlings from larger seeds begin life larger themselves, as has been extensively discussed in the literature \citep[reviewed by][]{Turnbull:2012ew}.
